---
title: "Motor Cars Data analysis"
subtitle: "Johns Hopkins Data Science Project: Regression Model"
author: "Suhas. P. K"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "dark"
    downcute_theme: "default"
---


```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
```

# Introduction
**Motor Trend** is a magazine about the automobile industry. In this analysis, a data set having information of car collection is explored to understand the relationship between **miles per gallon** and **transmission type**.\
The data set is available from the CRAN repository.\

## Libraries used.
```{r libraries, echo=FALSE}
library(ggplot2)
library(ggdark)
```


## Reading the data set.
```{r Read dataset mtcars}
data(mtcars)
```
### Basics checking of the data set.
```{r summary of mtcars}
summary(mtcars)
```
### Check the structure of data set.
```{r structure of mtcars}
str(mtcars)
```
### Check for the missing data.
```{r missing data in mtcars}
colSums(is.na(mtcars))
```
There are no missing value.

## Exploratory Data Analysis

Understand the distribution of **miles per gallon** variable.

```{r histogram of mpg}
histplot <- ggplot(data = mtcars,
                   aes(x = mtcars$mpg)) + geom_histogram(color = "black",fill = "lightgreen") +
    xlab("miles per gallon")+
    ggtitle("Histogram of Miles per gallon")+
    dark_theme_light()
histplot
```

Plotting a scatter plot based on the transmission type and mpg.
```{r scatter plot}
scttrplot <- ggplot(data = mtcars,
                    aes(x = am, y = mpg, color = factor(am)))
scttrplot + 
    geom_point(size = 2) + 
    geom_smooth(method=lm, color = "yellow")+ 
    dark_theme_light()
```






























